---
title: "001_data_prep"
author: "Eunjoo Byeon"
date: "5/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


# Data Wrangling
## Import

```{r}
hc_path = 'data/HC'
sz_path = 'data/SZ'

# getting a list of all txt file
hc_files = list.files(hc_path, 'txt', full.names = T)
sz_files = list.files(sz_path, 'txt', full.names = T)
```


Since overall length of information may not be consistent throughout, and the file iteself does not contain subject information, I'll run through one at a time.

Each stat file (txt) from dissection contains multiple (2 to 4, eventually 4) track groups. 
Each track groups (must) contain below 10 categories in 11 lines as below.
|  TrackGroup, Track count, Voxel Count, Mean values (Length, FA_Trk, Angtrk, FA, AD, MD, RD).


```{r}
# creating a dummy dataframe to fill
df <- data_frame(condition = "NONE", subnum = '', track_group = '', track_count = '', voxel_count = '', mean_length = '', mean_FAtrk = '', mean_Angtrk = '', mean_FA = '', mean_AD = '', mean_MD = '', mean_RD = '', error = '')
```

### TODO ###
add a script to match the subNum from filename to the FA/MD file reference name

```{r}
# Iterate over each txt file
process_txt <- function(filelist, conditionGroup, dataframe){
  for (file in filelist){
    path <- str_c(getwd(), '/', file) # file to read
    subNum <- str_extract(file, 'A000\\d\\d\\d\\d\\d') # extracting subject name
    txtfile <- file(path, 'r') # establishing connection 
    print(str_c('running subject ', subNum, '...'))
    while(TRUE){
      line <- readLines(txtfile, n=1) # read one line
      if (length(line) == 0){ # end of the file, close and exit
          close (txtfile)
          break
      } else { 
        if (str_detect(line, "TrackGroup")){ # make sure this is the first line of the track group
          trackGroup <- sub('TrackGroup: ', "", line)
          lines <- readLines(txtfile, n=10) # read the following 10 lines
          lines <- gsub(" ", "", lines) # since I want to extract variable names, remove all whitespace
          # store all relevant values
          for (i in 1:length(lines)){
            if (i != 3){ # this line has no value
              if (i %in% seq(7,10)){ # these are the lines with FA, MD, RD, AD values. Orders are not consistent, so we figure out the variable names by extracting the last two words before the value.
                if (str_extract(lines[i], 'A000\\d\\d\\d\\d\\d') != subNum){ # but first check if this is the right data (sanity check for administrative error)
                  err = TRUE
                } else {
                  err = FALSE
                }
                varName <- paste(str_sub(lines[i], 62, 63))
              } else {
                varName <- paste(sub(':.*', '', lines[i])) # extracting variable names
              }
            assign(varName, sub('.*:', '', lines[i])) # assigning values to the variables
            }
          }
          # end of one trackgroup, add a row
            dataframe <- add_row(condition = conditionGroup, subnum = subNum, track_group = trackGroup, track_count = Trackcount, voxel_count = Voxelcount, mean_length = Length, mean_FAtrk = FA_Trk, mean_Angtrk = AngTrk, mean_FA = FA, mean_AD = AD, mean_MD = MD, mean_RD = RD, error = err, dataframe)
        }  
      }
    }
  } 
  dataframe
}


df <- process_txt(hc_files, 'HC', df) # adding healthy controls
df <- process_txt(sz_files, 'SZ', df) # adding patient groups
df <- df[-1,] # removing a dummy row
head(df)
tail(df)
```

### Checking for errors
```{r}
# ones that filename don't match the data 
df %>% filter(error == TRUE) 
df <- df %>% select(-error)
```

```{r}
# duplicate data or multiple subjects
df %>% group_by(subnum, track_group) %>% summarise(n = n()) %>% filter(n > 1)
```



## Importing behavioral data

There are 3 sets of behavioral data.
SZ: 
schizconnect_COBRE_assessmentData_10541.csv

HC:
schizconnect_COBRE_assessmentData_10762.csv

COBRE_Data_Dictionary.xlsx : contains key


```{r}
sz_behav <- read_csv('data/SZ/Assessment/schizconnect_COBRE_assessmentData_10541.csv')
hc_behav <- read_csv('data/HC/Assessment/schizconnect_COBRE_assessmentData_10762.csv')

library('readxl')
behavKey <- read_excel('data/SZ/Assessment/COBRE_Data_Dictionary.xlsx')

```


### Cleaning behavioral data

```{r}
sz_behav_df <- sz_behav %>% 
  select(subjectid, assessment, assessment_description, question_id, question_value) %>%
  mutate(condition="SZ")
hc_behav_df <- hc_behav %>% 
  select(subjectid, assessment, assessment_description, question_id, question_value) %>% 
  mutate(condition = "HC")

sz_behav_df$assessment <- str_extract(sz_behav_df$assessment, pattern = "(?<=>)(.*?)(?=>)") # Extracting only the assessment type

hc_behav_df$assessment <- str_extract(hc_behav_df$assessment, pattern = "(?<=>)(.*?)(?=>)") 

names(behavKey) <- c("instrument", "q_label", "q_desc", "question_id", "r_label", "r_val", "r_desc") # cleaning up the column name

Key1 <- behavKey %>% select (instrument, q_label, question_id) # extracting key values

sz_behav_df <- left_join(sz_behav_df, Key1)
hc_behav_df <- left_join(hc_behav_df, Key1)
combined_behav <- rbind(sz_behav_df, hc_behav_df) # combining two conditions and adding key info

```

Spread the dataframe (pivot) so each measure is on its own column

```{r}
full_behav <- combined_behav %>% select(subjectid, condition, question_id, question_value)
full_behav <- full_behav %>% filter(!duplicated(full_behav)) # removing duplicates

# seems like some medication logs include 2 values for the sake of this analyses, I will average the two for now. (I filter this value out later, since I'm not using it for my analyses. But just in case.)

tmp <- full_behav %>% group_by(subjectid, question_id) %>% mutate(row = row_number()) %>% filter(row != 1)  # Select the question_id that has more than 1 row per participant
troublemakers <- unique(tmp$question_id) 
mean_scores <- full_behav %>% # averaging out the two values
  filter(question_id %in% troublemakers) %>% 
  filter (!is.na(question_value)) %>% 
  group_by (subjectid, question_id) %>% 
  summarise(question_value = mean(question_value)) 


full_behav<- full_behav %>% filter(!question_id %in% troublemakers) # replace those value
full_behav <- bind_rows(full_behav, mean_scores)

```

Before we pivot the dataframe, let's change question ids to something that's a bit more self-explanatory.

```{r}
Keys <- data.frame(question_id = unique(full_behav$question_id)) #creating key data frame
temp <- left_join(Keys, behavKey)

#I'm going to only look at behavioral task scores, demographic info, and PANSS scale. Edit this code if you want something else.

temp <- temp %>% 
  mutate(label = ifelse(startsWith(question_id, "CNP"), q_label, ifelse(startsWith(question_id, "CODEM"), str_c("DEMO-", q_label), ifelse(startsWith(question_id, "FIPAN"), str_c("PANSS-", q_label), NA)))) %>%
  select(question_id, label) %>% 
  filter(!is.na(label))

demo <- temp %>% #selecting only relevant demographic information (Age, Handedness, Gender, Ethnicity, Race, Marital Status, Highest level of education)
  filter(question_id %in% c("CODEM_1", "CODEM_10", "CODEM_2", "CODEM_3", "CODEM_3A", "CODEM_4", "CODEM_5"))

demo$label <- c("Age", "Handedness", "Gender", "Ethnicity", "Race", "Marital_St", "Highest_Education") # making label a bit easy to use
temp <- temp %>% filter(!startsWith(label, "DEMO")) # remove the rest of demographic information
temp$label <- gsub(" ", "", temp$label) # since I want to extract variable names, remove all whitespace

Keys <- rbind(temp, demo)
full_behav <- left_join(full_behav, Keys) %>% select(-question_id)
head(full_behav, 10)
```
```{r}
#value 9999 or 8888 is not applicable or missing, so change to that.
full_behav <- full_behav %>% mutate(question_value = ifelse(question_value %in% c(8888, 9999), NA, question_value))

```


Finally pivoting the dataframe... 
```{r}
sub_behav <- full_behav %>% filter(!is.na(label)) %>% pivot_wider(names_from = label, values_from = question_value)
head(sub_behav, 10)          

```

Unfortunately, it seems like some information is just not there for anybody If that's the case, drop that column.

```{r}
sub_behav <- sub_behav[!sapply(sub_behav, function(x) all(is.na(x)))]
names(sub_behav)
# unfortunately it seemed to have dropped most of demographic data
# maybe we should have run this sooner. 
```


### sanity check
Checking number of subjects in each conditions for behavioral data

```{r}
sub_behav %>% group_by(condition) %>% summarise(counts = n())
```


## Merging behav data with imaging data

```{r}

fulldf <- left_join(df, sub_behav, by = c("subnum" = "subjectid", "condition"))
glimpse(fulldf)

```




# next problems to fix..
## Change track_group names to be consistent
```{r}
unique(fulldf$track_group)
fulldf <- fulldf %>% 
  mutate(track_group = ifelse(track_group %in% c('Left', 'FATL', 'FAT_L', 'IFGL-SFGL', 'Aslant_L', 'Aslant_:'), 'FATL', ifelse(track_group %in% c('Right', 'FATR', 'FAT_R', 'IFGR-SFGR', 'Aslant_R'), 'FATR', ifelse(track_group %in% c('UF_R', 'UF__R'), 'UF_R', track_group)))) %>%
  mutate(track_group = as.factor(as.character(track_group)))

fulldf %>% group_by(condition, track_group) %>% summarise(count = n())
```

## Change number measures to be numbers 

First, I need to separate out mean and SEM

```{r}
fulldf <- fulldf %>% mutate(track_count = as.integer(track_count), voxel_count = as.integer(voxel_count)) # change counts as integers

#I'll write a quick function to separate mean and SEM 
SEM_log <- function(x){
  x <- str_remove_all(x, "[[:alpha:]]") #remove all texts (e.g. "mm")
  res <- list(mean = str_split(x, "\\+/-")[[1]][1], SEM = str_split(x, "\\+/-")[[1]][2]) # number before +/- sign is mean and the latter is SEM
  return (res)
}

varlist <- c('mean_length', 'mean_FAtrk', 'mean_Angtrk', 'mean_FA', 'mean_AD', 'mean_MD', 'mean_RD') # columns to run through this function

for (x in varlist){
  print(str_c('running...', x))
  semname <- paste(x, '_SEM', sep="")
  X <-fulldf[[x]]
  fulldf[[x]] <- unlist(t(sapply(1:length(X), function(x){SEM_log(X[x])}))[,1])
  fulldf[[semname]] <- unlist(t(sapply(1:length(X), function(x){SEM_log(X[x])}))[,2])
}


```

```{r}
semlist <- paste(varlist, "_SEM", sep="") # column list for SEMs
collist <- append(varlist, semlist) # colum list for means and SEMs

fulldf[collist] <- sapply(fulldf[collist], as.double) # make all of these double floats
fulldf[collist] <- fulldf[collist] %>% mutate_all(~replace(., is.na(.), 0)) # if NA they are 0 

summary(fulldf) # summary of final dataframe
```


# Export
Writing dataframe out as a Rdata file and a csv.
We'll use Rdata file in analyses. 

```{r}
write_csv(fulldf, "fulldf.csv")
save(fulldf, file = "fulldf.Rdata")
```


